---
title: "p8105_hw3_xm2276"
author: "XIAO MA"
date: "10/13/2021"
output: github_document
---

```{r}
#question is this leaflet necessary?
library(leaflet)
library(p8105.datasets)
data("instacart")
library(tidyverse)
library(dplyr)
library(ggridges)
```

There are **`r nrow(instacart)`**observations of the dataset.  
There are **`r ncol(instacart)` variables** of the dataset.  
The variables of the dataset are **`r colnames(instacart)`**.  
The **dimension of the dataset is `r dim(instacart)`**.  
The **length of the dataset is `r length(instacart)`**.
The key variables are 'reorded','order_hour_of_day','days_since_prior_order','product_name'
These variables are 

####describing some key variables, and giving illstrative examples of observations

problem 1
```{r}
janitor::clean_names(instacart)
aisle = instacart %>% 
  count(aisle) %>%
  arrange(desc(n))
##There are 134 aisles and 'Fresh Vegetables' are ordered most.
aisle_10000 = filter(aisle, n <= 10000)

```





Problem 2

##cleaning the data

```{r}
data("brfss_smart2010")
BRFSS = brfss_smart2010 %>%
  filter(Topic == "Overall Health") %>%
  mutate(
    Response = factor(Response, levels = c("Poor","Fair","Good","Very good", "Excellent")),
    Response = as.numeric(Response)) %>% 
    arrange(Response)
```

##Using this dataset, do or answer the following (commenting on the results of each):

##In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
year_2002 = filter(BRFSS, Year == '2002' ) %>%
  distinct(Locationabbr, Locationdesc) %>% 
   group_by(Locationabbr) %>%
   summarize(n_obs = n()) %>% 
   filter(n_obs >= 7)
year_2002
#by observing the dataset, there are 6 states observed at 7 or more locations. They are CT,FL,MA,NC,NJ,PA,
year_2010 = filter(BRFSS, Year == '2010' ) %>%
   distinct(Locationabbr, Locationdesc) %>% 
   group_by(Locationabbr) %>%
   summarize(n_obs = n()) %>% 
   filter(n_obs >= 7)
year_2010
#by observing the dataset, there are 14 states observed at 7 or more locations. They are CA,CO,FL,MA,MD,NC,NE,NJ,NY,OH,PA,SC,TX,WA

```

##Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r}
brfss_excellent = filter(BRFSS, Response == 5 ) %>%
   select(Year,Data_value, Locationabbr,Response) %>% 
   group_by(Year, Locationabbr) %>% 
   summarise_at(vars(Data_value), list(Average = mean))
brfss_excellent

#plot
ggplot(data = brfss_excellent, aes(x = Year, y = Average)) + geom_line(data = brfss_excellent,color = names())

p = 
  brfss_excellent %>% 
  ggplot(aes(x = Year, y = Average, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")


```

##Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}

```


problem 3
```{r}
accel_df = read_csv("data/accel_data.csv")

```

